Deriving an Executable Model of a Data Stream Protocol from its Specification

### 2009-09-15 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
paper ModelBasedTestCaseGenerationDataStreamProtocol: modified introduction added parts

METHODS: 
JML model.
JML tools to generate JUnit test suite.
JUnit test suite applied to developed system, to verify its correctness.
Junit test suite applied to simulator, to verify its correctness.

RESULTS: 
Proof developed system same behaviour as simulator.
Proof developed system and simulator verify the specifications.

@@

The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
sensor system installed at the University College Dublin in Dublin,
Ireland.  \textbf{Our facility provides a capability---unique in the world in
terms of its scale and flexibility---for performing in-depth
investigation of both the specific and general issues in large-scale
sensor networking.}

This system integrates a sensor platform, a datastore, and a compute
farm.  The sensor platform is a custom-designed, but inexpensive,
sensor platform (called the \ST) paired with general-purpose
small-scale compute nodes, including everything from low-power PDAs to
powerful touchscreen-based portable computers.  There are over a dozen
sensors on the \ST itself, and new sensors can be added by adding them
to the USB bus.

The \ST

\todo extend

@@

In section \ref{sec:data_stream_and_protocol_model} the data stream model 
and protocol is presented, we will give both a higl level overview of the 
protocol, and a formalized, JML based, specification,

In section \ref{sec:test_cases} the test cases structure is presented, for 
both the hand made and the generated ones. 
Both the approaches are analyzed, their pro and contro are discussed, as 
well as the expected advantages to use them in combination.

In section \ref{sec:test_cases} the test cases structure is presented, for 
both the hand made and the generated ones. 
Both the approaches are analyzed, their pro and contro are discussed, as 
well as the expected advantages to use them in combination.

@@

Finally in section \ref{sec:conclusions} we outline some conclusions and we 
briefly describe our ongoing and future works on similar topics.

@@

Finally, on top of the serial channels provided by the API, a custom protocol 
has been developed to manage the entire \STSB; that is, read sensor data 
observations, enable or disable sensors, configure onboard components 
as the Field-Programmable Gate Array (FPGA), and the micro processor; the 
protocol main function is to read sensor data.

@@

The internal structure of the output sensor data packet reflects the board 
capabilities and sensors.
Basically a single packet has to accomodate various types of data:
\begin{itemize}
 \item Fat data rate streams: audio data, 48KHz or 96KHz.
 \item Medium data rate streams: 5KHz.
 \item Slow data rate streams: temperature, pression, accellerometer, etc.; all 
 these sensor have a data rate far lower than 5 KHz.
 \item Metadatas: metadatas on sensors and board state.
\end{itemize}

Even if we have to deal with data streams that have different data rates, the 
internal structure of the package (as we will see) is strictly fixed. 
This choice originates from a tradeoff between two constraints: the efficiency 
of the trasmission (redundant data to be minimized) and the complexity to build 
a package (complexity to be minimized).
The complexity minimization constraint is simple to understand: it is the FPGA 
on the \STSB that build the packet, but it has also to be used for various 
other specific tasks. 
Hence, it is important to maintain the building packet code on the FPGA as 
simple and short as possible, to leave enough space for the code implementing 
other tasks.
The Output sensor data packet structure is showed in Table 
\ref{tab:data_packet_structure}, the packet total length is 1024 bytes (512 16 
bit words).

@@

The subsequence accomodate data from the fast and medium data rate streams 
(there are 4 fast data rate stream and 8 medium data rate stream channels). 
There are 82 subsequences in a packet.
The subsequence structure is shown in Table \ref{tab:subsequence_structure}.

@@

It is clear that, looking at the structure described and at the data rates 
frequencies, that there must be something subtle, to make all the figures to 
add up.
The subleties lies in two forms of ridundancies, for slow and mediumd data 
rates:

@@

\todo describe high level representation of package: sensors and data 
details remains, exact package structure and sensor data location in the
packege are lost

\todo describe JML specification of the high levelr representation of the 
package

@@

\todo benefit: the approch can be used to test different part of the system
that share the same specification

>>>>

METHODS: 
JML model.
JML tools to generate JUnit test suite.
JUnit test suite applied to developed system, to verify its correctness.
JUnit test suite applied to simulator, to verify its correctness.

RESULTS: 
Proof developed system same behavior as simulator.
Proof developed system and simulator verify the specifications.

@@

The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
sensor system to be installed at the University College Dublin in Dublin,
Ireland. 
Our facility aims to provide a scalable and extensible 
infrastructure for performing in-depth investigation of both the specific 
and general issues in large-scale sensor networking.

This system integrates a sensor platform, a datastore, and a compute
farm.  
The sensor platform is a custom-designed, but inexpensive,
sensor platform (called the \ST) paired with general-purpose
small-scale compute nodes, including everything from low-power PDAs to
powerful touchscreen-based portable computers. 
There are over a dozen
sensors on the \ST itself, and new sensors can be added by adding them
to the board.

Because of hard time constraints, and the initial unavailability of the 
custom board, we needed to progress concurrently with all the development tasks: 
the specification of the communication protocol with the /SB, the development of 
the controlling code in the /SB, the development of the driver to communicate 
with the /SB, and the development of the simulators needed for future tests 
of the system (the controlling code of the board was developed by the external 
manufacturer of the board, hence, it is not taken into account here).

\todo explain and refine the problem

\todo other approaches to the problem

To tackle the complexity of the concurrent tasks the formal model of the protocol 
was used as a focus point that lead all the other tasks. 
The formal model of the protocol was used to check at run time the correctness of 
the /SB, the driver, the simulator, and the hand made test suites of the driver 
and the simulator.
The formal model was used to automatically generate a test suite to verify that
both the driver and the simulator were compliant to the specification, and eventually 
to verify the specification itself.

The approach proved itself effective, at the point that, when the board was finally 
released, no defects were found in the driver or the simulator, while the errors and 
not agreed changes introduced in the board where discovered during the first execution.

@@

In section \ref{sec:data_stream_and_protocol_model} the data stream model 
and protocol is presented, we will present both a high level overview of the 
protocol, and a formalized, JML based, specification.

In section \ref{sec:test_cases} the test cases structure is presented, for 
both the hand made and the generated ones. 
Both the approaches are analyzed, their pro and contra are discussed, as 
well as the expected advantages to use them in combination.

@@

Finally in section \ref{sec:conclusions} conclusions are drawed and 
ongoing and future works on similar topics is introduced.

@@

Finally, on top of the serial channels provided by the API, a custom protocol 
has been developed to manage the entire \STSB; that is, read sensor data 
observations, enable or disable sensors, configure on-board components 
as the Field-Programmable Gate Array (FPGA), and the micro processor; the 
protocol main function is to read sensor data.

@@

The internal structure of the output sensor data packet reflects the board 
capabilities and sensors.
Basically a single packet has to accommodate various types of data:
\begin{itemize}
 \item Fast data rate streams: audio data, 48KHz or 96KHz.
 \item Medium data rate streams: 5KHz.
 \item Slow data rate streams: temperature, pression, accelerometer, etc.; all 
 these sensor have a data rate far lower than 5 KHz.
 \item Metadatas: metadatas on sensors and board state.
\end{itemize}

Even if we have to deal with data streams that have different data rates, the 
internal structure of the package (as we will see) is strictly fixed. 
This choice originates from a trade-off between two constraints: the efficiency 
of the transmission (redundant data to be minimized) and the complexity to build 
a package (complexity to be minimized).
The complexity minimization constraint is simple to understand: it is the FPGA 
on the \STSB that build the packet, but it has also to be used for various 
other specific tasks. 
Hence, it is important to maintain the building packet code on the FPGA as 
simple and short as possible, to leave enough space for the code implementing 
other tasks.
The Output sensor data packet structure is showed in Table 
\ref{tab:data_packet_structure}, the packet total length is 1024 bytes (512 16 
bit words).

@@

The subsequence accommodate data from the fast and medium data rate streams 
(there are 4 fast data rate stream and 8 medium data rate stream channels). 
There are 82 subsequences in a packet.
The subsequence structure is shown in Table \ref{tab:subsequence_structure}.

@@

Looking at the structure described and at the data rates frequencies, it
can be seen that there must be something subtle, to make all the figures to 
add up.
The subtleties lies in two forms of redundancies, for slow and medium data 
rates:

@@

\todo describe high level representation of package: sensors and data 
details remains, exact package structure and sensor data location in the
package are lost

\todo describe JML specification of the high level representation of the 
package

@@

\todo benefit: the approach can be used to test different part of the system
that share the same specification
